---
title: \vspace{-3.5cm}  Statistical Inference Course Project
author: "David Contreras Carretero"
output: 
  pdf_document:
    fig_width: 6
    fig_height: 3.5
---
\vspace{-0.5cm}  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
## Part 1: Simulation Exercise Instructions

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

 **1. Show the sample mean and compare it to the theoretical mean of the distribution.** 
 
First we create a thousand simulations of the 40 exponentials and calculate and store their means.
```{r}
set.seed(1928)
lambda = 0.2
n = 40
nsim=1000

sims <- data.frame(mean=numeric(nsim))

for (i in 1 : nsim) {
  sample <-  rexp(n,lambda)
  sims[i,1] <- mean(sample)
}
```
```{r}
tmean = 1/lambda
smean <-mean(sims$mean)
```
   Sample mean : `r smean` 
   Theoretical mean: `r tmean`
   
```{r echo=FALSE}
ggplot(sims, aes(x=mean)) +
           geom_histogram(binwidth=0.1) +
           labs(y = "Number of obs", 
                x = "Sample means", 
                title ="Figure 1 - Distribution of sample means of 40 exponentials",
                caption ="The blue line indicates theoretical mean, red line sample mean") +
           geom_vline(xintercept=tmean,  color = "blue", size=1) +
           geom_vline(xintercept=smean,  color = "red", size=1) 
```
  
   The histrogram show how close theoretical and sample mean are. 
  
  **2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.**
  
  We know that theoretical standard deviation is 1/lambda, so theoretical variance is 1/lambda^2 
  
```{r}
tvar <-  (1/lambda^2)/n
tvar
```
  
    Then we calculate de sample variance as the variance of the means.
```{r}
svar <- var(sims$mean)
svar
```
  As we can see theoretical variance and sample variance are also very close.
  
  **3. Show that the distribution is approximately normal.**
  
  In order to compare the distribution of averages of samples with a normal distribution with theoretical mean and standard deviation.
  
```{r echo = FALSE}
ggplot(sims, aes(x=mean)) +
       labs(y = "Number of obs", x = "Sample means", title =" Figure 2 - Distribution of sample means of 40 exponentials" ,caption ="Green line show distribution density, red lines show normal distribution and theoretical mean") +
      geom_histogram(aes(y=..density..), size=1, binwidth=0.2) + 
      geom_density(color="green", size=1) +
      stat_function(fun=dnorm, args=list(mean=tmean, sd=sqrt(tvar)), color = "red", size=1) +
      geom_vline(xintercept=tmean, color="red", size=1) 
```

  Figure 2 shows how sample distribution almost match  normal distribution line. We would say it is aproximatly normal. 
\newpage
## Appendix - R Code
```{r echo=TRUE, eval=FALSE}
# Figure 1 code
ggplot(sims, aes(x=mean)) +
           geom_histogram(binwidth=0.1) +
           labs(y = "Number of obs", 
                x = "Sample means", 
                title ="Figure 1 - Distribution of sample means of 40 exponentials",
                caption ="The blue line indicates theoretical mean, red line sample mean") +
           geom_vline(xintercept=tmean,  color = "blue", size=1) +
           geom_vline(xintercept=smean,  color = "red", size=1) 
```
```{r echo = TRUE, eval=FALSE}
# Figure 2 code

ggplot(sims, aes(x=mean)) +
       labs(y = "Number of obs",
            x = "Sample means", 
            title =" Figure 2 - Distribution of sample means of 40 exponentials",
            caption ="Green line show distribution density, 
                      red lines show normal distribution and theoretical mean") +
      geom_histogram(aes(y=..density..), size=1, binwidth=0.2) + 
      geom_density(color="green", size=1) +
      stat_function(fun=dnorm, args=list(mean=tmean, sd=sqrt(tvar)), color = "red", size=1) +
      geom_vline(xintercept=tmean, color="red", size=1) 
```
